---
authors: Carlos aybar , Biao Chen, Aarmando Mcghee
date: 04/12/19
title: "R Notebook"
output: html_notebook
---

## Orissa

```{r}
install.packages("party")
```

```{r}
library(tidyverse)
library(party)
```


```{r}
rainfall <- read.csv("rainfall.csv")
head(rainfall)
```


```{r}
orissa <- subset(rainfall , YEAR > 1915 & SUBDIVISION == "ORISSA")
head(orissa)
```


### Hyphothesis TESTING
# Exploratory analysis
Scatter plot of adjusted earnings per year.
```{r}
ggplot(data = orissa, 
       mapping = aes(x = YEAR, y = ANNUAL, color = YEAR)) +
  geom_line(show.legend = TRUE) 
```

# comparing orissa's rainfall from jun-sep with the rainfall from march - may
```{r}
t.test(orissa$Jan.Feb, orissa$Mar.May, paired = FALSE)

cat("p-value > alpha (0.05). The null hypothesis fails to be rejected.\n")
cat("The adjusted earnings for Trilogies 2 and 3 are NOT significantly different.\n")
```

```{r}
t.test(orissa$Jun.Sep, orissa$Oct.Dec, paired = FALSE)

cat("p-value > alpha (0.05). The null hypothesis fails to be rejected.\n")
cat("The rainfall during the months of jun-sep and oct-dec is not significantly differentt.\n")
```


## ANOVA
# Checking to see if there is a significant difference between all the mothns
ANOVA performs a significant difference test on means of 3 or more treatments.
ANOVA determines if there is a significant difference among the treatments.
The aov(<data>, <formula>) function performs the analysis.
  The <data> must be a single data frame with each of the rows labels as treatments.
  The <formula> parameter is the variable for the average ~ the treatment variable.
  Example: aov(dataframe, formula = attribute ~ treatment)

Is there a significant difference over the years in the annual rainfall in Orissa
```{r}
anovaTrilogy <- aov(data = orissa, ANNUAL ~ YEAR)
summary(anovaTrilogy)
```




## LINEAR REGRESSION
# Visualize the relationship between the data
rainfall during january-february vs march to may
```{r}
ggplot(data = orissa, mapping = aes(x = Mar.May, y = Jun.Sep)) +
  geom_point(mapping = aes(color = 2)) +
  geom_smooth(method = "lm") +
  labs(title = "Annual rainfal for in Orissa  year" ,
         subtitle = "rainfall from 1915-2015" ,
         x = "March to May")
```

### anlisisis comment : the more rainfall orissa gets during march and may, the more it will get 
### during jun and sep

finding the correlation in the annual rainfall over the years
```{r}
summerRain <- select(orissa, c("Mar.May", "Jun.Sep"))
pairs(summerRain)
```


## here we calculate the correlation
Calculate correlation value [-1, 1]
```{r}
rainfallCorrelation <- cor(orissa$Mar.May, orissa$Jun.Sep)
```

## Univariate Linear Regression
Given 1 independent variable, predict the dependent variable.
given the rainfall for march-may, we will predict the amount that orissa is expected to get
for jun-sep
```{r}
model <- lm(Jun.Sep ~ Mar.May, data = orissa)
model
```


## Prediction
Given a goal difference of 16, predict the position.
```{r}
# Create a data frame with a single GD entry
test <- data.frame(Mar.May = c(162))

# Predict the position with the model and test data
predictedPosition <- predict(model, test)
cat("when orissa gets a rainfall of 162 from march to may, they will get", predictedPosition ," from june to september")
```


## CLASSIFICATION
### reading the first 100 rows of the dataset orissa
```{r}
#Get the first 105 students
reading <- orissa[c(1:500),]
reading
```


## decision tree
### We are creating a decision tree to predict the annual rainfall in orissa 
```{r}

tree <- ctree(ANNUAL ~ MAY + JUN, data = orissa)
tree
```
```{r}
plot(tree)
```


### We are using the modelcreated above and the new one in order to predict the annual rainfall
### based on the two months with the highest rainfall in Orissa
```{r}
#predicting whether it is summer, fall, or spring, based on the rainfall
testSeason <- data.frame(MAY = as.double(c(495.1)),
                          JUN = as.double(c(368.7)))

#making the prediction based on the model
predictedSeason <-  predict(tree , newdata = testSeason)
cat("based on the specific rainfall for may and june, the annual rainfall will be : " , predictedSeason)
```


# installing the knitr and rmarkdown packages
```{r}
install.packages("knitr")
install.packages("rmarkdown")
install.packages("forecast")
```



# calling the installed pacages
```{r}
library(knitr)
library(rmarkdown)
library(forecast)
```


## TIME SERIES
### extracting the months and the years from the orissa data frame

```{r}
orissa2 <- orissa[, c(2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12 ,13, 14)]
print(orissa2)
```


# printing the new data set
```{r}
data(orissa2)
orissa2
```



# Plotting the new data set
```{r}
plot(orissa2)
```

### trainning the data
```{r}
# training set: 1949 to 1956
train = window(orissa2, start = 1920, end = c(1960, 12))
# test set:  1957 to 1960 
test = window(airSeries, start = 1961, end = c(2000, 12))
```


## clustering and verification
```{r}

library(ggplot2)

```

```{r}
head(orissa)
```


### we will be clustering the data into may, jun, and july because they are the months with most rain
```{r}
set.seed(20)
summer <- data.frame(Year = orissa$YEAR , June = orissa$JUN , July = orissa$JUL, August = orissa$AUG)
head(summer)

```

## K-Means: Perform Algorithm
```{r}
#perform kmeans with k set to 3
orissaCluster <- kmeans(summer , 3 , nstart = 25)
orissaCluster
```

## verification
```{r}
table(orissaCluster$cluster , orissa$YEAR)
```

## visualization
```{r}
#add the cluster assignment to each point
orissa$YEAR <- as.factor(orissaCluster$cluster)



#get centroids
centroids <- as.data.frame(orissaCluster$centers)
centroids$Cluster <- as.factor(c(1:3))


#visualize cluster assignmetns
ggplot(data = summer , aes(June, July , August, color = Year)) +
  geom_point() +
  geom_point(data = centroids , aes(x = June, y = July, fill = Cluster) , size = 5, shape =13)+
  labs(titel = "Rainfall June" , x = " Rainfall July" , y = "Width (cm)")
```



### wthin sum of Squares (WSS)
```{r}
#for each k, perform WSS, store the value
wss <- numeric(15)
for(k in 1:15)
  wss[k] = sum(kmeans(summer, k , nstart = 25) $withinss)
```

```{r}
#Make a data frame out of the WSS result
wssResults <- data.frame(k = c(1:15), wss = wss)
wssResults
```


```{r}
ggplot(data  = wssResults  , aes(x = k  , y = wss))+
  geom_point( )+
  geom_line() +
  labs("k-means: iris" , x = "number of Clusters k" , y = "within sum of squares")
 
```




```{r}
orissa1 <- subset(rainfall , YEAR > 2004 & SUBDIVISION == "ORISSA")
month<-c(orissa1$JAN,orissa1$FEB,orissa1$MAR,orissa1$APR,orissa1$MAY,orissa1$JUN,orissa1$JUL,orissa1$AUG,orissa1$SEP,orissa1$OCT,orissa1$NOV,orissa1$DEC)

monthts<-ts(month,frequency=12,start = c(2005))
plot(monthts)
```

```{r}
library(knitr)
library(rmarkdown)
library(forecast)
```



```{r}
# training set: 1949 to 1956
train = window(monthts, start = 2004, end = c(2010, 12))
# test set:  1957 to 1960 
test = window(monthts, start =2011, end = c(2015, 12))

arimaModel <- auto.arima(train)
arimaForecast <-forecast(arimaModel, h = 60)
arimaForecast
```

```{r}
plot(arimaForecast, main = "rain", ylab = "rainfall", xlab = "Months")
lines(test, col="red")
legend("topleft",lty = 1,bty = "n",col=c("red","blue"),c("Actual","Predicted"))
```

```{r}
arimaMean <-arimaForecast$mean
plot(test, main = "rain", ylab = "rainfall", xlab = "Months", col = "red")  
lines(arimaMean, col="blue")
legend("topleft",lty = 1,bty = "n",col=c("red","blue"),c("Actual","Predicted"))
```






```{r}
install.packages("party")
```

```{r}
library(tidyverse)
library(party)
```


```{r}
rainfall <- read.csv("rainfall.csv")
View(rainfall)
```


## KERALA
	
	
```{r}
KERALA <- subset(rainfall , YEAR > 1915 & SUBDIVISION == "KERALA")
head(KERALA)
```


```{r}
ggplot(data = KERALA, 
       mapping = aes(x = YEAR, y = ANNUAL, color = YEAR)) +
  geom_line(show.legend = TRUE) 
```

```{r}
t.test(KERALA$Jan.Feb, KERALA$Mar.May, paired = FALSE)

cat("p-value > alpha (0.05). The null hypothesis fails to be rejected.\n")
cat("The adjusted earnings for Trilogies 2 and 3 are NOT significantly different.\n")
```






## ANOVA
### Checking to see if there is a significant difference between all the mothns

ANOVA performs a significant difference test on means of 3 or more treatments.
ANOVA determines if there is a significant difference among the treatments.
The aov(<data>, <formula>) function performs the analysis.
  The <data> must be a single data frame with each of the rows labels as treatments.
  The <formula> parameter is the variable for the average ~ the treatment variable.
  Example: aov(dataframe, formula = attribute ~ treatment)

Is there a significant difference over the years in the annual rainfall in Orissa
```{r}
anovaTrilogy1 <- aov(data = KERALA, ANNUAL ~ YEAR)
summary(anovaTrilogy1)
```



## LINEAR REGRESSION
# Visualize the relationship between the data
rainfall during january-february vs march to may
```{r}
ggplot(data = KERALA, mapping = aes(x = Mar.May, y = Jun.Sep)) +
  geom_point(mapping = aes(color = 2)) +
  geom_smooth(method = "lm") +
  labs(title = "Annual rainfal for in Orissa  year" ,
         subtitle = "rainfall from 1915-2015" ,
         x = "March to May")
```

### anlisisis comment : the more rainfall orissa gets during march and may, the more it will get 
### during jun and sep

finding the correlation in the annual rainfall over the years
```{r}
summerRain1 <- select(KERALA, c("Mar.May", "Jun.Sep"))
pairs(summerRain1)
```


### here we calculate the correlation
Calculate correlation value [-1, 1]
```{r}
rainfallCorrelation3 <- cor(KERALA$Mar.May, KERALA$Jun.Sep)
```

### Univariate Linear Regression
Given 1 independent variable, predict the dependent variable.
given the rainfall for march-may, we will predict the amount that orissa is expected to get
for jun-sep
```{r}
model3 <- lm(Jun.Sep ~ Mar.May, data =KERALA)
model3
```


## Prediction
Given a goal difference of 16, predict the position.
```{r}
# Create a data frame with a single GD entry
test3 <- data.frame(Mar.May = c(162))

# Predict the position with the model and test data
predictedPosition <- predict(model3, test3)
cat("when orissa gets a rainfall of 162 from march to may, they will get", predictedPosition ," from june to september")
```















## CLASSIFICATION
### reading the first 100 rows of the dataset orissa
```{r}
#Get the first 105 students
reading3 <- KERALA[c(1:500),]
reading3
```


## decision tree
### We are creating a decision tree to predict the annual rainfall in orissa 
```{r}

tree3 <- ctree(ANNUAL ~ MAY + JUN, data = KERALA)
tree3
```
```{r}
plot(tree3)
```


### We are using the modelcreated above and the new one in order to predict the annual rainfall
### based on the two months with the highest rainfall in Orissa
```{r}
#predicting whether it is summer, fall, or spring, based on the rainfall
testSeason3 <- data.frame(MAY = as.double(c(251.0)),
                          JUN = as.double(c(454.4)))

#making the prediction based on the model
predictedSeason3 <-  predict(tree3 , newdata = testSeason3)
cat("based on the specific rainfall for may and june, the annual rainfall will be : " , predictedSeason3)
```



# installing the knitr and rmarkdown packages
```{r}
install.packages("knitr")
install.packages("rmarkdown")
install.packages("forecast")
```



## clustering and verification
```{r}

library(ggplot2)

```



### we will be clustering the data into may, jun, and july because they are the months with most rain
```{r}
set.seed(20)
summer3 <- data.frame(Year = KERALA$YEAR , June = KERALA$JUN , July = KERALA$JUL, August = KERALA$AUG)
head(summer3)

```

## K-Means: Perform Algorithm
```{r}
#perform kmeans with k set to 3
KERALACluster <- kmeans(summer3 , 3 , nstart = 25)
KERALACluster
```

## verification
```{r}
table(KERALACluster$cluster , KERALA$YEAR)
```

## visualization
```{r}
#add the cluster assignment to each point
KERALA$YEAR <- as.factor(KERALACluster$cluster)



#get centroids
centroids3 <- as.data.frame(KERALACluster$centers)
centroids3$Cluster <- as.factor(c(1:3))


#visualize cluster assignmetns
ggplot(data = summer3 , aes(June, July , August, color = Year)) +
  geom_point() +
  geom_point(data = centroids3 , aes(x = June, y = July, fill = Cluster) , size = 5, shape =13)+
  labs(titel = "Rainfall June" , x = " Rainfall July" , y = "Width (cm)")
```



## wthin sum of Squares (WSS)
```{r}
#for each k, perform WSS, store the value
wss3 <- numeric(15)
for(k in 1:15)
  wss3[k] = sum(kmeans(summer3, k , nstart = 25) $withinss)
```

```{r}
#Make a data frame out of the WSS result
wssResults3 <- data.frame(k = c(1:15), wss = wss3)
wssResults3
```


```{r}
ggplot(data  = wssResults3  , aes(x = k  , y = wss3))+
  geom_point( )+
  geom_line() +
  labs("k-means: iris" , x = "number of Clusters k" , y = "within sum of squares")
 
```




```{r}
KERALA3 <- subset(rainfall , YEAR > 2004 & SUBDIVISION == "KERALA")
month<-c(KERALA3$JAN,KERALA3$FEB,KERALA3$MAR,KERALA3$APR,KERALA3$MAY,KERALA3$JUN,KERALA3$JUL,KERALA3$AUG,KERALA3$SEP,KERALA3$OCT,KERALA3$NOV,KERALA3$DEC)

monthts3<-ts(month,frequency=12,start = c(2005))
plot(monthts3)
```

```{r}
library(knitr)
library(rmarkdown)
library(forecast)
```



```{r}
# training set: 1949 to 1956
train3 = window(monthts3, start = 2004, end = c(2009, 12))
# test set:  1957 to 1960 
test3 = window(monthts3, start =2010, end = c(2011, 12))

arimaModel3 <- auto.arima(train3)
arimaForecast3 <-forecast(arimaModel3, h = 24)
arimaForecast3
```

```{r}
plot(arimaForecast3, main = "rain", ylab = "rainfall", xlab = "Months")
lines(test3, col="red")
legend("topleft",lty = 1,bty = "n",col=c("red","blue"),c("Actual","Predicted"))
```

```{r}
arimaMean3 <-arimaForecast3$mean
plot(test3, main = "rain", ylab = "rainfall", xlab = "Months", col = "red")  
lines(arimaMean3, col="blue")
legend("topleft",lty = 1,bty = "n",col=c("red","blue"),c("Actual","Predicted"))
```